---
title: "STAT/MATH 495: Advanced Data Analysis"
author: "Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_depth: 1
    toc_float: true
---

<style>
h1{font-weight: 400;}
</style>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning = FALSE)
library(tidyverse)
library(broom)
library(knitr)
library(modelr)
library(lubridate)
library(forcats)
```

```{r, eval=FALSE, echo=FALSE}
# Run this separately to have slide output:
rmarkdown::render("index.Rmd", output_format = c("ioslides_presentation"), output_file = "slides.html")
```



***



# X.X

## Thu X/X

Announcements:

* Problem set 0X feedback and example solutions given [below](#PS0X).


## Chalk talk

X.X




***



# X.X 

## Tue X/X

Announcements:

* Problem set 0X presentations listed [below](#PS0X).




***


# Problem set 0X {#PS0X}

## Information

*Assigned on Tue X/X, due on Tue X/X 9am at which point there will be synchronized pull requests.*

You will...

* Due: Tue X/X 9:00am, at which point there will be synchronized pull requests.
* Teams: 
* Deliverables: One ready to submit *pull request* to the [PS0X repo](https://github.com/2017-09-Amherst-STAT495/PS0X) (recall the GitHub [submission process](https://github.com/2017-09-Amherst-STAT495/README#problem-set-submission-process)). At the very least, the following files should be modified/added:
    1. `PS0X.Rmd`: This should
    1. `PS0X.html`: The output of `PS0X.Rmd`. This will be the presentation file for a randomly chosen set of groups.
    1. `README.md`: Change `rudeboybert` in the URL to the GitHub ID of the team leader so that you can open your team's version of `PSX2.html` in a browser on Tuesday morning.
* Tips:
* Hints:


## Presentations


## Feedback


## Example solutions





***



# X.X

## Mon X/X

Announcements:



## Chalk talk

X.X



***



