---
title: "ROC Curves Shiny App"
author: "Albert Y. Kim"
date: "Last updated on `r Sys.Date()`"
output: html_document
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


## Background

```{r, echo=FALSE, message=FALSE}
values <- read_csv("https://raw.githubusercontent.com/rudeboybert/MATH218/gh-pages/assets/classification/ROC_shiny/example.csv")
```

The first three of the `r nrow(values)` rows of input data look like:

```{r, echo=FALSE, message=FALSE}
values %>% 
  head(3) %>% 
  knitr::kable()
```

**Rule**: If $\widehat{p} > p^*$, then predict $\widehat{y} = 1$.



## ROC Curve as a Function of $p^*$

```{r eruptions, echo=FALSE}
inputPanel(
  sliderInput("p_star", label = "Prob Threshold p* (Red Dot):", min = 0, max = 1, value = 0.5, step = 0.01)
)

renderPlot({
  ROC_data <- values %>% 
    group_by(p_hat) %>%
    summarise(
      Positive = sum(y),
      Negative = n() - sum(y)
      ) %>%
    arrange(-p_hat) %>%
    mutate(
      TPR = cumsum(Positive) / sum(Positive),
      FPR = cumsum(Negative) / sum(Negative)
      )
  
  AUC <- ROC_data %>% 
    summarise(AUC = sum(diff(FPR) * na.omit(lead(TPR) + TPR)) / 2)

    current <- ROC_data %>% 
    mutate(
      diff_p = abs(p_hat - input$p_star)
    ) %>% 
    arrange(diff_p) %>% 
    slice(1)
  
  ggplot(ROC_data, aes(FPR, TPR)) +
    geom_line() +
    geom_abline(lty = 2) +
    labs(
      x = expression(paste("False Positive Rate(", p, "*)")),
      y = expression(paste("True Positive Rate(", p, "*)")), 
      title = paste("ROC Curve as a Parametric Function of p*. Area Under Curve = ", round(AUC,3), sep="")
      ) +
    geom_point(data=current, size=8, col="red")
})

renderTable({
  if(input$p_star == 0){
    output <- 
      data_frame(
        truth = as.factor(c(0, 0, 1, 1)),
        predicted = as.factor(c(0, 1, 0, 1)),
        n = as.integer(c(0, sum(values$y == 0), 0, sum(values$y == 1)))
      )
  } else if(input$p_star == 1){
    output <- 
      data_frame(
        truth = as.factor(c(0, 0, 1, 1)),
        predicted = as.factor(c(0, 1, 0, 1)),
        n = as.integer(c(sum(values$y == 0), 0, sum(values$y == 1), 0))
      )
  } else {
    output <- values %>% 
      mutate(y_hat = ifelse(p_hat>input$p_star, 1, 0)) %>% 
      select(y, y_hat) %>% 
      rename(truth=y, predicted=y_hat) %>% 
      mutate(
        truth = as.factor(truth),
        predicted=as.factor(predicted)
      ) %>% 
      group_by(truth, predicted) %>% 
      count()
  }
  blah <- output$n %>% matrix(nrow=2) %>% as.data.frame()
  rownames(blah) <- c("<strong>Predict y=0</strong>", "<strong>Predict y=1</strong>")
  colnames(blah) <- c("Truth: y=0", "Truth: y=1")
  blah
}, rownames = TRUE, sanitize.text.function=function(x){x})


renderUI({
  withMathJax()
  
  if(input$p_star == 0){
    output <- 
      data_frame(
        truth = as.factor(c(0, 0, 1, 1)),
        predicted = as.factor(c(0, 1, 0, 1)),
        n = as.integer(c(0, sum(values$y == 0), 0, sum(values$y == 1)))
      )
  } else if(input$p_star == 1){
    output <- 
      data_frame(
        truth = as.factor(c(0, 0, 1, 1)),
        predicted = as.factor(c(0, 1, 0, 1)),
        n = as.integer(c(sum(values$y == 0), 0, sum(values$y == 1), 0))
      )
  } else {
    output <- values %>% 
      mutate(y_hat = ifelse(p_hat>input$p_star, 1, 0)) %>% 
      select(y, y_hat) %>% 
      rename(truth=y, predicted=y_hat) %>% 
      mutate(
        truth = as.factor(truth),
        predicted=as.factor(predicted)
      ) %>% 
      group_by(truth, predicted) %>% 
      count()
  }
  
  TN <- output$n[1]
  FP <- output$n[2]
  FN <- output$n[3]
  TP <- output$n[4]
  
  TPR <- TP/(TP+FN)
  FPR <- FP/(FP+TN)

str1 <- paste("False Positive Rate = ", FP, "/(", FP, "+", TN, ") = ", round(FPR,3), sep="")
str2 <- paste( "True Positive Rate = ", TP, "/(", TP, "+", FN, ") = ", round(TPR,3), sep="")
str <- paste( "$$\\mbox{False Positive Rate} = \\frac{", FP, "}{", FP, "+", TN, "} = \\thinspace ", round(FPR,3), "\\mbox{ & }","\\thinspace \\mbox{True Positive Rate} = \\frac{", TP, "}{", TP, "+", FN, "} = ", round(TPR,3),"$$", sep="")
HTML(paste(str1, str2, sep = '<br/>'))
  
})
```


## Note

An ROC Curve does *not* plot a function $(x, f(x))$, but rather a 
*parametric curve* $(x(\alpha), y(\alpha))$ as a function of parameter $\alpha$ where:

* $\alpha$: Probability threshold $p^*$
* $x(\alpha)$: Resulting False Positive Rate $\mbox{FPR}(p^*)$
* $y(\alpha)$: Resulting True Positive Rate $\mbox{TPR}(p^*)$
